---
title: "Midterm"
author: "Akhila Nekkanti"
date: "4/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)

library(glue)
library(tidyverse)
library(ggtext)
library(gapr)

```


# Part A: Data 

## 1. Use the function to download all the data from 2014-15 to each of the past 4 school years and bind it into a single data frame, using a single function (i.e., one line of code). Note, this may take a minute or two to run, depending on your Internet speed.

```{r download_file}

download_file <- function(year) {
    link <- glue::glue("https://www.oregon.gov/ode/educator-resources/assessment/TestResults20{year}/pagr_schools_ela_raceethnicity_{year-1}{year}.xlsx")
    rio::import(link, setclass = "tibble", na = c("-", "--", "*"))
}

years <- 15:18

# loop through years, apply download file function
data <- map_df(years, ~download_file(.x)) %>% 
  janitor::clean_names()


```

## Data Cleaning
```{r data cleaning}

td <- data %>% 
  pivot_longer(contains("number_level"), 
  names_to = "level", 
  values_to = "n") %>%  
  mutate(level = str_extract(level, "[0-9]")) %>% 
  select(academic_year, 
         district, school, 
         student_group, 
         grade_level, 
         level, 
         n) %>% 
  filter(student_group == "Hispanic/Latino" |
           student_group == "White") %>% 
  drop_na()
  
```

##2. (5 points) Collapse (sum) the n across grades and school for each district within each academic year, student group, and level. At the same time, calculate the number of schools (unique schools listed) for each district. Also move your data to a wider format where the student groups become new columns, filled by the district n. Remove districts that did not report numbers for one or both student groups (i.e., drop missing data from the new columns). Your data frame should look like the below

```{r}

#df1 has all the unique schools in each district
unique_school <- td %>% 
  group_by(district) %>% 
  summarise(n_schools = length(unique(school))) %>% 
  ungroup()

#df2 has the number of students in each district, by academic year and level
students <- td %>% 
  group_by(district, academic_year, student_group, level) %>% 
  summarise(n = sum(n)) %>% 
  pivot_wider(names_from = student_group, 
              values_from = n) %>% 
  janitor::clean_names() %>% 
  ungroup()

#join the 2 df together, drop NA, re-order the columns
td2 <- left_join(students, unique_school) %>% 
  drop_na() %>% 
  select(academic_year,
         district,
         n_schools,
         level,
         hispanic_latino,
         white) %>% 
  mutate(academic_year = as.factor(academic_year))

td2

```
# Part B: Acheivement Gaps 

```{r}

#estimate_v estimates the average diference between two groups

td[1:4, ] %>% # data source - just first four rows for first district
  estimate_v("white", "hispanic_latino") # columns for reference/focal groups

```

##1. (10 points) Estimate an achievement gap effect size for every school district in the state that reported data on both student groups (i.e., using the data we created above), for each academic year. Note, on my computer this took about 38 seconds. It will likely take a bit longer for you, but shouldn’t take too long. You may get a warning. Just go ahead and ignore that for now.

```{r}

#nest data by year and district 
by_district <- td2 %>%
    nest(-district, -academic_year, -n_schools)


#loop estimate_v function through each district, for each academic year by using the nested df above
results_by_district <- by_district %>%
  mutate(gap = map(by_district$data, ~estimate_v("white", "hispanic_latino", 
                                  data = .x))) %>% 
  select(-data) %>% 
  unnest(gap)


```


##2. (10 points) The plot below shows the achievement gap estimate for one school district across years. Produce a similar plot to the below for each of the first 100 school districts. Make sure you don’t miss the caption noting the number of schools in the district.

Hint: You need to select unique districts.

```{r}

#find the first 100 unique districts from our results -- character
dists <- unique(results_by_district$district)[1:100]

#create new object that selects for the 100 we identified above
pd <- results_by_district %>% 
  filter(district %in% dists)

#check
unique(pd$district)


```

## First try with a subset.

```{r}

#create subset to test code
pd_subset <- pd %>% 
  filter(district == "Adrian SD 61" |
           district == "Amity SD 4J")

# try on one plot
#call as.numeric on academic year so geom_line can connect dots 
pd %>% filter(district == "Bandon SD 54") %>% 
  ggplot(aes(academic_year, v)) + 
  geom_point(aes(color = v > 0)) +
  geom_line(aes(x = as.numeric(academic_year))) +
  geom_errorbar(aes(ymin = v - v_se, ymax = v + v_se))


#loop through subset of data
psub <- pd_subset %>% 
  group_by(district) %>% 
  nest() %>% 
  mutate(plots = pmap(list(district, data),
                ~ggplot(..2, aes(academic_year, v)) +
                        geom_point(aes(color = v > 0)) +
                        scale_colour_manual(values =  c("#A563CA", "#6DA3F9")) +
  geom_line(aes(x = as.numeric(academic_year)), , color = "grey70") +
  geom_errorbar(aes(ymin = v - v_se, ymax = v + v_se), color = "grey70") +
  geom_hline(yintercept = 0, color = "#cb615d"))) 

psub$plots[2]

```

## Final Plots
```{r}

final_plots <- pd %>%
  group_by(district, n_schools) %>% 
  nest () %>% 
    mutate(plots = pmap(list(district, n_schools, data),
    ~ggplot(..3, aes(academic_year, v)) +
      geom_point(aes(color = v > 0), size = 2) +
          scale_color_manual(values = c("#C468F7", "#6DA3F9")) +
      geom_line(aes(x = as.numeric(academic_year)),
              color = "grey70") +
      geom_errorbar(aes(ymin = v - v_se, ymax = v + v_se),color = "grey70") +
      geom_hline(yintercept = 0, color = "red", size = 0.5) +
      geom_point(aes(color = v > 0)) +
          scale_color_manual(values = c("#C468F7", "#6DA3F9")) + #added twice, because removing this from above layer gives an error
      theme(plot.title.position = "plot",
          plot.subtitle = element_markdown(),
          legend.position = "none",
          ) +
    labs(x = "School year",
         y = "Average achievement difference (SD units)",
         caption = glue("Note: District included {..2} schools."),
         title = glue("Achievement Gap Estimates: {..1}"), 
         subtitle = "Negative values indicate <span style = 'color:#C468F7;'>White</span> students have higher average achievement  
       Positive values indicate <span style = 'color:#6DA3F9;'>Hispanic/Latino</span> students have higher average achievement.")))


# looking at the plots: 
final_plots$plots[[8]]


```


## 3. (10 points) Save the plots into a “plots” directory. Make sure the file names are meaningful.
```{r}

fs::dir_create(here::here("akhila_plots"))

files <- str_replace_all(tolower(final_plots$district), " ", "_")
paths <- here::here("akhila_plots", glue("{files}.png"))
paths

walk2(paths, final_plots$plots, ggsave,
       width = 9.5, 
       height = 6.5,
       dpi = 500)


```

