---
title: "midterm_joanna"
author: "Joanna Wright"
date: "4/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)

# install.packages("glue")
library(glue)
library(tidyverse)

```

# Part A: Data 

##1. Use the function to download all the data from 2014-15 to each of the past 4 school years and bind it into a single data frame, using a single function (i.e., one line of code). Note, this may take a minute or two to run, depending on your Internet speed.

```{r}

download_file <- function(year) {
    link <- glue::glue("https://www.oregon.gov/ode/educator-resources/assessment/TestResults20{year}/pagr_schools_ela_raceethnicity_{year-1}{year}.xlsx")
    rio::import(link, setclass = "tibble", na = c("-", "--", "*"))
}

```

```{r}
# sy15 <- download_file(15)

years <- 15:18

# loop through years, apply download file function
data <- map_df(years, ~download_file(.x))

#OR this also works: 
# map_df(years, download_file)

```


Conduct some basic data cleaning to make your data file look like the following.

Filter for only student groups coded as "White" or "Hispanic/Latino".
Select variables related to the number of students in each of the levels (1:4), and not percentages or collapsed levels.
Remove any row that has missing data in any of the n variables

```{r data cleaning}

data1 <- data %>%
  pivot_longer(
    cols = c(11, 13, 15, 17),
    names_to = "level", 
    values_to = "n")
# need to remember how to use names_pattern to drop part of string (change "Number Level #" to just "#") 

data1 <- data1 %>%
  select(`Academic Year`, District, School, `Student Group`, `Grade Level`, level, n) %>% 
  filter(`Student Group` == "Hispanic/Latino" |
           `Student Group` == "White") %>%
  filter(n != "NA")

```

##2. (5 points) Collapse (sum) the n across grades and school for each district within each academic year, student group, and level. At the same time, calulate the number of schools (unique schools listed) for each district. Also move your data to a wider format where the student groups become new columns, filled by the district n. Remove districts that did not report numbers for one or both student groups (i.e., drop missing data from the new columns). Your data frame should look like the below

```{r}

data1 %>%
  group_by(District, `Academic Year`, `Student Group`, level) %>%
  summarise(n = sum(n)) %>%
  pivot_wider(
    names_from = `Student Group`, 
    values_from = n)

data1 %>%
  group_by(District, level, `Academic Year`) %>%
  summarise(n_distinct(School))

```

#Part B: Acheivement Gaps 

```{r}
library(gapr)

```

##1. (10 points) Estimate an achievement gap effect size for every school district in the state that reported data on both student groups (i.e., using the data we created above), for each academic year. Note, on my computer this took about 38 seconds. It will likely take a bit longer for you, but shouldn’t take too long. You may get a warning. Just go ahead and ignore that for now.

```{r}

```


##2. (10 points) The plot below shows the achievement gap estimate for one school district across years. Produce a similar plot to the below for each of the first 100 school districts. Make sure you don’t miss the caption noting the number of schools in the district.

Hint: You need to select unique districts.

```{r}

```

##3. (10 points) Save the plots into a “plots” directory. Make sure the file names are meaningful.

```{r}

```
