---
title: "Rowley_Mid-Term"
author: "Brock Rowley"
date: "4/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)

library(glue)
library(tidyverse)
library(gapr)
library(ggtext)
```

# Part A: Data
## 1. Use the function to download all the data from 2014-15 to each of the past 4 school years and bind it into a single data frame, using a single function (i.e., one line of code).
```{r download}
download_file <- function(year) {
    link <- glue::glue("https://www.oregon.gov/ode/educator-resources/assessment/TestResults20{year}/pagr_schools_ela_raceethnicity_{year-1}{year}.xlsx")
    rio::import(link, setclass = "tibble", na = c("-", "--", "*"))
}
years <- 15:18
data <- map_df(years, ~download_file(.x))

#Could be written as:
# map_df(years, download_file)
```

```{r cleaning}
names(data)

td <- data %>%
  pivot_longer(contains("Number Level"),
  names_to = "level", 
  values_to = "n") %>%  
  mutate(level = str_extract(level, "[0-9]")) %>% 
  select(`Academic Year`, District, School, `Student Group`, `Grade Level`, level, n) %>% 
  filter(`Student Group` == "Hispanic/Latino" |
           `Student Group` == "White") %>% 
  drop_na()
  
td <- td %>%
  janitor::clean_names()

td
```

## 2. (5 points) Collapse (sum) the n across grades and school for each district within each academic year, student group, and level. At the same time, calulate the number of schools (unique schools listed) for each district. Also move your data to a wider format where the student groups become new columns, filled by the district n. Remove districts that did not report numbers for one or both student groups (i.e., drop missing data from the new columns). Your data frame should look like the below
```{r clean2}
# Two df's joined them together with unique schools.
unique_school <- td %>%
  group_by(district) %>%
  summarise(n_schools = length(unique(school))) %>%
  ungroup()

# Get students (n)
students <- td %>%
  group_by(district, academic_year, student_group, level) %>%
  summarise(n = sum(n)) %>%
  pivot_wider(names_from = student_group,
              values_from = n) %>%
  janitor::clean_names() %>%
  ungroup()

# Two df's joined together, drop NA, re-order the columns.
td2 <- left_join(students, unique_school) %>%
  drop_na() %>%
  select(academic_year,
         district,
         n_schools,
         level,
         hispanic_latino,
         white
         )
```

# Part B: Acheivement Gaps
## 1. (10 points) Estimate an achievement gap effect size for every school district in the state that reported data on both student groups (i.e., using the data we created above), for each academic year. Note, on my computer this took about 38 seconds. It will likely take a bit longer for you, but shouldn’t take too long. You may get a warning. Just go ahead and ignore that for now.
```{r gaps}
# First four rows for first district, columns for groups.
td2[1:4, ] %>%
  gapr::estimate_v("white", "hispanic_latino")

# Academic year and district as factor.
td2 <- td2 %>%
  mutate(academic_year = as.factor(academic_year))

d_split <- split(td2, td2$district)

# Do for two school districts.
sub_split <- d_split[1:2]

# This returns a factor.
sub_split[[1]]$academic_year

# This returns a tibble.
sub_split[[1]][1]

# Nest by district and year.
by_district <- td2 %>%
    nest(-district, -academic_year, -n_schools)

# Checking with only two districts
sub_by_district_year <- td2 %>%
  filter(district == "Adrian SD 61" |
           district == "Amity SD 4J") %>%
    nest(-district, -academic_year)

sub_by_district_year_results <- sub_by_district_year %>%
  mutate(gap = map(sub_by_district_year$data, ~estimate_v("white", "hispanic_latino",
                                  data = .x))) %>%
           select(-data) %>%
           unnest(gap)

# Moving to entire data.
results_by_district <- by_district %>%
    mutate(gap = map(by_district$data, ~estimate_v("white", "hispanic_latino",
                                                   data = .x))) %>%
           select(-data) %>%
           unnest(gap)
```

## 2. (10 points) The plot below shows the achievement gap estimate for one school district across years. Produce a similar plot to the below for each of the first 100 school districts. Make sure you don’t miss the caption noting the number of schools in the district.

Hint: You need to select unique districts.

```{r unique}
dists <- unique(results_by_district$district)[1:100]

pd <- results_by_district %>%
  filter(district %in% dists)

unique(pd$district)

# Number of rows.
pd %>% count(district)

pd_subset <- pd %>%
  filter(district == "Adrian SD 61" |
           district == "Amity SD 4J")

# Help from D.A. here: as.factor() makes acacdemic year a factor; then call as.numeric() to make academic_year numeric in order to get geam_line to work.
pd %>% filter(district == "Bandon SD 54") %>%
  ggplot(aes(academic_year, v)) +
  geom_point(aes(color = v > 0)) +
  geom_line(aes(x = as.numeric(as.factor(academic_year)))) +
  geom_errorbar(aes(ymin = v - v_se, ymax = v + v_se)) +
  geom_hline(yintercept = 0, color = "#CC615D") +
  labs(x = "School year",
       y = "Average achievement difference (SD units)")

# Trying with one (Bandon) school district.
pd2 <- pd %>%
    mutate(label =
    glue("Note. District included {(n_schools)} schools."))

tmp_pd <- pd2 %>%
  filter(district == "Bandon SD 54")

ggplot(tmp_pd, aes(academic_year, v)) +
  geom_point(aes(color = v > 0)) +
  geom_line(aes(x = as.numeric(as.factor(academic_year)))) +
  geom_errorbar(aes(ymin = v - v_se, ymax = v + v_se)) +
  geom_hline(yintercept = 0, color = "#CC615D") +
  theme(
    plot.title.position = "plot",
    plot.subtitle = element_markdown()) +
  labs(x = "School year",
       y = "Average achievement difference (SD units)",
       title = glue("Achievement Gap Estimates: {(tmp_pd$district)}"),
       subtitle = "Negative values indicate <span style = 'color:#C468F7;'>White</span> students have higher average achievement  
       Positive values indicate <span style = 'color:#6DA3F9;'>Hispanic/Latino</span> students have higher average achievement.",
       caption = unique(tmp_pd$label))

# Nest data to loop through lists for all plots.
pd2 %>%
    group_by(district, label) %>%
    nest()

final_plots <- pd2 %>%
    group_by(district, label) %>%
    nest() %>%
    mutate(plots = pmap(list(district, label, data),
    ~ggplot(..3, aes(academic_year, v), show.legend = FALSE) +
        geom_point(aes(color = v > 0), show.legend = FALSE) +
    geom_line(aes(x = as.numeric(as.factor(academic_year)))) +
    geom_errorbar(aes(ymin = v - v_se, ymax = v + v_se)) +
    geom_hline(yintercept = 0, color = "#CC615D") +
    theme(
    plot.title.position = "plot",
    plot.subtitle = element_markdown()) +
  labs(x = "School year",
       y = "Average achievement difference (SD units)",
       title = glue("Achievement Gap Estimates: {(pd2$district)}"),
       subtitle = "Negative values indicate <span style = 'color:#C468F7;'>White</span> students have higher average achievement  
       Positive values indicate <span style = 'color:#6DA3F9;'>Hispanic/Latino</span> students have higher average achievement.",
       caption = unique(pd2$label))))

# Look at the plots:
final_plots$plots[[1]]
```

## 3. (10 points) Save the plots into a “plots” directory. Make sure the file names are meaningful.

```{r save_plots}

```
