---
title: "EDLD 610: Functional Programming Midterm"
author: "Stephanie Gluck"
date: "5/3/2020"
output: 
  html_document: 
    highlight: kate
    theme: paper
---

```{r setup}
knitr::opts_chunk$set(cache = TRUE)

library(tidyverse)
library(ggtext)
library(janitor)
library(glue)
```


## Part A: Data 

### 1. Use the function to download all the data from 2014-15 to each of the past 4 school years and bind it into a single data frame, using a single function (i.e., one line of code). Note, this may take a minute or two to run, depending on your Internet speed.

```{r download_file}

download_file <- function(year) {
    link <- glue::glue("https://www.oregon.gov/ode/educator-resources/assessment/TestResults20{year}/pagr_schools_ela_raceethnicity_{year-1}{year}.xlsx")
    rio::import(link, setclass = "tibble", na = c("-", "--", "*"))
}

```

```{r download_data}
years <- 15:18

# loop through years, apply download file function
data <- map_df(years, ~download_file(.x)) %>% 
  janitor::clean_names()

#OR this also works: 
# map_df(years, download_file)

```

Conduct some basic data cleaning to make your data file look like the following.

Filter for only student groups coded as "White" or "Hispanic/Latino".
Select variables related to the number of students in each of the levels (1:4), and not percentages or collapsed levels.
Remove any row that has missing data in any of the n variables

```{r data_cleaning}

names(data)

td <- data %>% 
  pivot_longer(contains("number_level"), 
  names_to = "level", 
  values_to = "n") %>%  
  mutate(level = str_extract(level, "[0-9]")) %>% 
  select(academic_year, district, school, student_group, grade_level, level, n) %>% 
  filter(student_group == "Hispanic/Latino" |
           student_group == "White") %>% 
  drop_na()
  
td

```

### 2. (5 points) Collapse (sum) the n across grades and school for each district within each academic year, student group, and level. At the same time, calulate the number of schools (unique schools listed) for each district. Also move your data to a wider format where the student groups become new columns, filled by the district n. Remove districts that did not report numbers for one or both student groups (i.e., drop missing data from the new columns). Your data frame should look like the below

```{r data_wrangling}

str(td)
names(td)

#create 2 separate df and join them together 
#unique schools 

unique_school <- td %>% 
  group_by(district) %>% 
  summarise(n_schools = length(unique(school))) %>% 
  ungroup()

#number of students (n)
students <- td %>% 
  group_by(district, academic_year, student_group, level) %>% 
  summarise(n = sum(n)) %>% 
  pivot_wider(names_from = student_group, 
              values_from = n) %>% 
  janitor::clean_names() %>% 
  ungroup()

#join the 2 df together, drop NA, re-order the columns
td2 <- left_join(students, unique_school) %>% 
  drop_na() %>% 
  select(academic_year, district, n_schools, level, hispanic_latino, white)

td2

```

## Part B: Acheivement Gaps 

### 1. (10 points) Estimate an achievement gap effect size for every school district in the state that reported data on both student groups (i.e., using the data we created above), for each academic year. Note, on my computer this took about 38 seconds. It will likely take a bit longer for you, but shouldn’t take too long. You may get a warning. Just go ahead and ignore that for now.

```{r effect_size}

library(gapr)

#nest data by year and district 
by_district <- td2 %>%
    nest(-district, -academic_year, -n_schools)
by_district

#run achievement gap effect size and extract auc, v and v_se
results_by_district <- by_district %>%
  mutate(gap = map(data, ~estimate_v("white", "hispanic_latino", 
                                  data = .x))) %>% 
  select(-data) %>% 
  unnest(gap)
```


### 2. (10 points) The plot below shows the achievement gap estimate for one school district across years. Produce a similar plot to the below for each of the first 100 school districts. Make sure you don’t miss the caption noting the number of schools in the district.

Hint: You need to select unique districts.

```{r unique_districts}
dists <- unique(results_by_district$district)[1:100]

pd <- results_by_district %>% 
  filter(district %in% dists)

unique(pd$district)

#to examine number of rows 
pd %>% count(district)

```

```{r plot,  fig.height = 6, fig.width = 8}

theme_set(theme_minimal(14))

#create one plot first
# as.factor() makes acacdemic year a factor; then can call as.numeric() on this to make academic_year numeric in order to connect via geam_line: 
pd %>% filter(district == "Bandon SD 54") %>% 
  ggplot(aes(academic_year, v)) + 
  geom_point(aes(color = v > 0)) +
  geom_line(aes(x = as.numeric(as.factor(academic_year)))) +
  geom_errorbar(aes(ymin = v - v_se, ymax = v + v_se)) +
  geom_hline(yintercept = 0, color = "red") +
  labs(x = "School year",
       y = "Average achievement difference (SD units)")


#create column that is the caption we want: 
pd2 <- pd %>%
    mutate(label = 
    glue("Note. District included {(n_schools)} schools."),
    title = 
    glue("Achievment Gap Estimates: {(district)}"))

#create multiple plot
pd2 <- pd2 %>%
    group_by(district, label) %>%
    nest()

final_plots <- pd2 %>%
    mutate(plots = pmap(list(district, label, data),
    ~ggplot(..3, aes(academic_year, v)) +
      geom_point(aes(color = v > 0)) +
    geom_line(aes(x = as.numeric(as.factor(academic_year))),
              color = "grey70", 
              size = 1) +
    geom_errorbar(aes(ymin = v - v_se, ymax = v + v_se),
                  color = "grey70", 
                  size = 1, 
                  width = 0.4) +
            geom_point(aes(color = v > 0), 
                       size = 3) +
      scale_color_manual(values = c("#C468F7", "#6DA3F9")) +
    geom_hline(yintercept = 0, color = "red",
               size = 1) +
    theme(plot.title.position = "plot",
          plot.subtitle = element_markdown(),
          legend.position = "none") +
    labs(x = "School year",
         y = "Average achievement difference (SD units)",
         caption = glue("{..2}"),
         title = glue("Achievement Gap Estimates: {..1}"), 
         subtitle = "Negative values indicate <span style = 'color:#C468F7;'>White</span> students have higher average achievement  
       Positive values indicate <span style = 'color:#6DA3F9;'>Hispanic/Latino</span> students have higher average achievement.")))


# looking at one of the plots: 
final_plots$plots[[8]]


```


### 3. (10 points) Save the plots into a “plots” directory. Make sure the file names are meaningful.

```{r save_plots}


fs::dir_create(here::here("stephanie_plots"))

# trying for first 2 districts to make sure it works: 
tmp_final_plots <- final_plots %>%
  filter(district == "Adrian SD 61" |
           district == "Amity SD 4J")

tmp_final_plots$plots[[2]]

files <- str_replace_all(tolower(tmp_final_plots$district), " ", "_")
paths <- here::here("stephanie_plots", glue("{files}.png"))

# paths
walk2(paths, tmp_final_plots$plots, ggsave,
       width = 9, 
       height = 6,
       dpi = 100)

# generalizing to all files:

files <- str_replace_all(tolower(final_plots$district), " ", "_")
paths <- here::here("stephanie_plots", glue("{files}.png"))
# paths
walk2(paths, final_plots$plots, ggsave,
       width = 9, 
       height = 6,
       dpi = 100)

```

