---
title: "Stephanie Midterm"
author: "Stephanie Gluck"
date: "4/22/2020"
output: html_document
---

https://uo-datasci-specialization.github.io/c3-fun_program2020/midterm.html

```{r setup}
knitr::opts_chunk$set(cache = TRUE)

library(tidyverse)
library(janitor)
```


# Part A: Data 

##1. Use the function to download all the data from 2014-15 to each of the past 4 school years and bind it into a single data frame, using a single function (i.e., one line of code). Note, this may take a minute or two to run, depending on your Internet speed.

```{r download_file}

download_file <- function(year) {
    link <- glue::glue("https://www.oregon.gov/ode/educator-resources/assessment/TestResults20{year}/pagr_schools_ela_raceethnicity_{year-1}{year}.xlsx")
    rio::import(link, setclass = "tibble", na = c("-", "--", "*"))
}

```

```{r download_data}
years <- 15:18

# loop through years, apply download file function
data <- map_df(years, ~download_file(.x)) %>% 
  janitor::clean_names()

#OR this also works: 
# map_df(years, download_file)

```

Conduct some basic data cleaning to make your data file look like the following.

Filter for only student groups coded as "White" or "Hispanic/Latino".
Select variables related to the number of students in each of the levels (1:4), and not percentages or collapsed levels.
Remove any row that has missing data in any of the n variables

```{r data cleaning}

names(data)

td <- data %>% 
  pivot_longer(contains("number_level"), 
  names_to = "level", 
  values_to = "n") %>%  
  mutate(level = str_extract(level, "[0-9]")) %>% 
  select(academic_year, district, school, student_group, grade_level, level, n) %>% 
  filter(student_group == "Hispanic/Latino" |
           student_group == "White") %>% 
  drop_na()
  
td

```

##2. (5 points) Collapse (sum) the n across grades and school for each district within each academic year, student group, and level. At the same time, calulate the number of schools (unique schools listed) for each district. Also move your data to a wider format where the student groups become new columns, filled by the district n. Remove districts that did not report numbers for one or both student groups (i.e., drop missing data from the new columns). Your data frame should look like the below

```{r}

str(td)
names(td)

#create 2 separate df and join them together 
#unique schools 

unique_school <- td %>% 
  group_by(district) %>% 
  summarise(n_schools = length(unique(school))) %>% 
  ungroup()

#number of students (n)
students <- td %>% 
  group_by(district, academic_year, student_group, level) %>% 
  summarise(n = sum(n)) %>% 
  pivot_wider(names_from = student_group, 
              values_from = n) %>% 
  janitor::clean_names() %>% 
  ungroup()

#join the 2 df together, drop NA, re-order the columns
td2 <- left_join(students, unique_school) %>% 
  drop_na() %>% 
  select(academic_year, district, n_schools, level, hispanic_latino, white)

td2

```

#Part B: Acheivement Gaps 

```{r}
library(gapr)

```

##1. (10 points) Estimate an achievement gap effect size for every school district in the state that reported data on both student groups (i.e., using the data we created above), for each academic year. Note, on my computer this took about 38 seconds. It will likely take a bit longer for you, but shouldn’t take too long. You may get a warning. Just go ahead and ignore that for now.

```{r}

#nest data by year and district 
by_district <- td2 %>%
    nest(-district, -academic_year, -n_schools)
by_district

#run achievement gap effect size and extract auc, v and v_se
results_by_district <- by_district %>%
  mutate(gap = map(data, ~estimate_v("white", "hispanic_latino", 
                                  data = .x))) %>% 
  select(-data) %>% 
  unnest(gap)
```


##2. (10 points) The plot below shows the achievement gap estimate for one school district across years. Produce a similar plot to the below for each of the first 100 school districts. Make sure you don’t miss the caption noting the number of schools in the district.

Hint: You need to select unique districts.

```{r}
dists <- unique(results_by_district$district)[1:100]

pd <- results_by_district %>% 
  filter(district %in% dists)

unique(pd$district)

#to get number of rows 
pd %>% count(district)

```

##3. (10 points) Save the plots into a “plots” directory. Make sure the file names are meaningful.

```{r}

pd_subset <- pd %>% 
  filter(district == "Adrian SD 61" |
           district == "Amity SD 4J")


#need to define year as numeric (from character) in order to connect line
pd %>% filter(district == "Bandon SD 54") %>% 
  ggplot(aes(academic_year, v)) + 
  geom_point() +
  geom_line(aes(x = as.numeric(as.factor(academic_year)))) +
  geom_errorbar(aes(ymin = v - v_se, ymax = v + v_se))



```




```{r, eval = FALSE}
td2[1:4, ] %>% # data source - just first four rows for first district
  gapr::estimate_v("white", "hispanic_latino") 

#make academic year & district a factor 
td2 <- td2 %>% 
  mutate(academic_year = as.factor(academic_year))

str(td2)

d_split <- split(td2, td2$district)

#map estimate_v for all school district 
#gap <- map(d_split, ~estimate_v("white", "hispanic_latino", data = .x))
#yay worked! (took a while to run)

#extract only 2 school district to play w code
sub_split <- d_split[1:2]
str(sub_split)

#returns factor 
sub_split[[1]]$academic_year

#returns tibble
sub_split[[1]][1]

#works for across school district (not separated by year)
gap <- map(sub_split, ~estimate_v("white", "hispanic_latino", 
                                  data = .x))
gap



#subset by only 2 district to simplify and for playing w code
sub_by_district <- td2 %>%
  filter(district == "Adrian SD 61" |
           district == "Amity SD 4J") %>% 
    nest(-district)
    
sub_by_district


# by_district_gap_sub <- sub_by_district %>%
#   mutate(year = map(data, "academic_year"))
   
str(by_district_gap_sub)

dists <- unique(by_district_gap$district)[1:100]

test <- by_district_gap %>% 
  filter(district %in% dists)

test$gap

str(test$gap)

#need to extract v and v_se

test[["gap"]][[1]]["v"]
test[["gap"]][[1]]["v_se"]

#use unnest to extract v and v_se
test <- test %>% 
  mutate(tidied = map(gap, broom::tidy)) %>% 
  select(academic_year, district, tidied) %>%  
  unnest() %>%  
  select(academic_year, district, column, mean)
  

str(test)

test <- test %>% pivot_wider(names_from = "column", values_from = "mean")
test

#testing codes not working 
test %>%
    mutate(v = map_dbl(gap, ~([[.x]][[1]]["v"])))

p12 = map_dbl(comp12, list("Pr(>F)", 2)))

#extract v and v_se
test$gap[[1]]$v

str(test2)

test2$v
test$gap


```

